# Group Assignment - Group 2

# **Created by Anastasia Kampantai (s253724), Hannah Espelund Nørrelykke (s224972), Mie He Larsen (s224976), Raminta Jonauskaite (s254123)**

## Load Libraries

```{r}
#| message: false

library(tidyverse)
library("broom")
library(cowplot)
library("ggrepel")
```

## Load Data

In the assignment we use the data from the paper "**A prognostic DNA signature for T1T2 node-negative breast cancer patients"** related to gene expression.

```{r}
raw_dir <- "data/_raw/"
data_file <- "gravier.RData"
data_loc <- "https://github.com/ramhiser/datamicroarray/raw/master/data/"
if( !dir.exists(raw_dir) ){
  dir.create(path = raw_dir)
}
if( !file.exists(str_c(raw_dir, data_file)) ){
  download.file(
    url = str_c(data_loc, data_file),
    destfile = str_c(raw_dir, data_file))
}

load(file = str_c(raw_dir, data_file))
```

## Clean Data

```{r}
gravier_clean_aug <- gravier |>
  bind_cols() |>
  as_tibble() 
```

## PCA

```{r}
gravier_PCA <- gravier_clean_aug |> 
  select(where(is.numeric)) |> 
  prcomp(scale = TRUE)
```

## **Mie He Larsen (s224976)**

## **Hannah Espelund Nørrelykke (s224972)**

For information to be clear from the plot 15 random genes are chosen

```{r}
set.seed(3)

gravier_PCA |> 
  tidy(matrix = "rotation") |> 
  pivot_wider(names_from = "PC", 
              names_prefix = "PC", 
              values_from = "value") |>
  slice_sample(n = 15) |> 
  ggplot(aes(PC1, PC2)) +
  geom_segment(xend = 0,
               yend = 0,
               alpha = 0.25) +
  geom_text_repel(aes(label = column))
```

## **Raminta Jonauskaite (s254123)**

## **Anastasia Kampantai (s253724)**
