---
title: "Group Document"
format: html
editor: visual
---

## Anastasia Kampantai

This is some code from assignment 6, added for the completion of this task.

Here is the code chunk for loading the gravier data.

```{r}
library(ggrepel)
library(tidyverse)

raw_dir <- "data/_raw/"
data_file <- "gravier.RData"
data_loc <- "https://github.com/ramhiser/datamicroarray/raw/master/data/"
if( !dir.exists(raw_dir) ){
  dir.create(path = raw_dir)
}
if( !file.exists(str_c(raw_dir, data_file)) ){
  download.file(
    url = str_c(data_loc, data_file),
    destfile = str_c(raw_dir, data_file))
}

load(file = str_c(raw_dir, data_file))

```

### New Analysis

This is a new section for the task. Also, from Assignment6 to clean data.

```{r}
gravier_clean_aug <- gravier |>
  bind_cols() |>
  as_tibble() |> 
   mutate(y = case_when(y == "poor" ~ 1,
                       y == "good" ~ 0)) |> 
  relocate(early_metastasis = y)
```

## Hannah Nørrelykke

Just some code

```{r}
c <- 3
1 + c
```

### New analysis

Mean and sd for random data

```{r}
data_H <- rnorm(n = 100,
                mean = 5,
                sd = 3)
mean(data_H)
sd(data_H)
```

## Mie He Larsen

#### New feature

```{r}
x1 <- 4
x2 <- 2
x1 + x2
```

## Raminta Jonauskaite

This is a code for running tidyverse

```{r}
library(tidyverse)
```

## New analysis

This is a code from previous exercises

```{r}
ggplot(mpg, aes(displ, hwy, color = drv)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  ggrepel::geom_label_repel(data = label_info, aes(label = drive_type), size = 5, nudge_y = 2) +
  theme(legend.position = "none")

```
